Contextualizing Climate Ensembles

### 2016-06-24 exam2.tex
mostly just reference revision
updated abstract & stuff based on Tufte's talk...

\title{Visualizing Relationships Amongst and With Climate Ensembles}
\author{Hannah Aizenman}
\committee{ Dr. Michael Grossberg(Advisor), Robert Haralick, Huy Vo}
\submitted{}
\abstract{Climate datasets often inherently try to encapsulate complex systems. The variables are often high dimensional and often have long ranging interactions both temporally and spatially. Ensemble climate models often add another layer to the complexity because they add uncertainty to the already complex interactions. Scientists often try to visualize these interactions either by attempting to encode many if them in a visualization, using a machine learning algorithm to model the interaction and traditional approaches to visualize them, or building out a complex toolkit that relies on users to compare graphs of each of the sub interactions to build out the relationships. Tamara'Munzner's domain-task-idiom framework can be an effective way of grouping the visualization techniques that have become dominant in the ensemble evaluation community and understanding what it is they're trying to accomplish.}
\begin{document}
\makefrontmatter

@@

>>>>

\title{Contextualizing Climate Ensembles}
\author{Hannah Aizenman}
\committee{ Dr. Michael Grossberg(Advisor), Robert Haralick, Huy Vo}
\submitted{}
\abstract{
  Problem: Ensembles inherently encode uncertainity since they are different
  variations on the same model. While there are many ways to visualize
  ensembles, and the uncertianity in the ensembles, these visualizations are
  often at least a step or two removed from what is happening in the real
  world. 
  Importance: Climate & weather ensembles are widley used by scientists and
  policy makers to evaluate risk; this is readily seen in Hurrican season when
  governor's have to decide on whether to move people based on how sure they
  are that the model's predictions are really accurate. Decontextualized
  visualizations that are more about the method than the data can occlude this
  decision process.

  Solution: THere are a kitchen sinks worth of methods and toolkits for
  visualizing ensembles, but the best ones clearly articulate the
  spatio-temporal relationships in the ensembles and between the variables such
  that the user can definitively parse out what a line or dot is saying about
  the temperature in New Jersey.  
}

\begin{document}
\makefrontmatter

@@

``Contextualize all the things'' is important. Tufte stresses it because he's
confirmatory, but Munzner's exploratray task-data-idiom approach is also
grounded in it. This is because while very pretty visualizations look well
pretty, at the end of the day they're semi-useless if there are too many steps
removed between the visualization and the underlying data.
